---
title: "Lake Sunapee Protective Association Buoy DO Visual History"
author: "B. Steele, Weathers Lab, Cary Institute"
date: "02 March 2022"
output: word_document
---

```{r setup, include=FALSE}
#set global options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

#load libraries
library(tidyverse)
library(ggthemes)

#store theme grom for ggplot
final_theme=theme_bw() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title=element_text(size=16, face='bold', hjust=0.5)) #save as a grom

```


```{r buoy data, include=FALSE}
datadir = 'C:/Users/steeleb/Dropbox/Lake Sunapee/monitoring/buoy data/data/all sensors/L1/do/'
exodir = 'C:/Users/steeleb/Dropbox/Lake Sunapee/monitoring/buoy data/data/all sensors/L1/exo/'

do2007 <- read.csv(file.path(datadir, '2007_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location) %>% 
  filter(buoy_location == 'loon')
do2008 <- read.csv(file.path(datadir, '2008_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location) %>% 
  filter(buoy_location == 'loon')
do2009 <- read.csv(file.path(datadir, '2009_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2010 <- read.csv(file.path(datadir, '2010_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2011 <- read.csv(file.path(datadir, '2011_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2012 <- read.csv(file.path(datadir, '2012_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2013 <- read.csv(file.path(datadir, '2013_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2014 <- read.csv(file.path(datadir, '2014_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2015 <- read.csv(file.path(datadir, '2015_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2016 <- read.csv(file.path(datadir, '2016_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2017 <- read.csv(file.path(datadir, '2017_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2018 <- read.csv(file.path(datadir, '2018_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2019 <- read.csv(file.path(datadir, '2019_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2020 <- read.csv(file.path(datadir, '2020_do_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2021lower <- read.csv(file.path(datadir, '2021_do_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')
do2021upper <- read.csv(file.path(exodir, '2021_exo_L1_v2022.csv')) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')

do2018hobo <- read.csv(file.path(datadir, '2018_hobodo_L1_v2022.csv'))  %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'),
         data_source = 'buoy') %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log'))) %>% 
  rename(buoy_location = location)%>% 
  filter(buoy_location == 'loon')

```


```{r manual do measurements, include=FALSE}
mando_dir = 'C:/Users/steeleb/Dropbox/Lake Sunapee/monitoring/buoy data/data/all sensors/L1/manual_do/'
mando = read.csv(file.path(mando_dir, 'manual_do_2007-2021_v2022-02-25.csv')) %>% 
  rename(site = location) %>% 
  mutate(datetime = as.POSIXct(date)) %>% 
  mutate(year = format(as.Date(date), '%Y')) %>% 
  mutate(data_source = factor(data_source, levels = c('buoy', 'LMP', 'operations log')))

mando_2007 <-  mando %>% 
  filter(year == 2007)

mando_2008 <-  mando %>% 
  filter(year == 2008)

mando_2009 <-  mando %>% 
  filter(year == 2009)

mando_2010 <-  mando %>% 
  filter(year == 2010)

mando_2011 <-  mando %>% 
  filter(year == 2011)

mando_2012 <-  mando %>% 
  filter(year == 2012)

mando_2013 <-  mando %>% 
  filter(year == 2013)

mando_2014 <-  mando %>% 
  filter(year == 2014)

mando_2015 <-  mando %>% 
  filter(year == 2015)

mando_2016 <-  mando %>% 
  filter(year == 2016)

mando_2017 <-  mando %>% 
  filter(year == 2017)

mando_2018 <-  mando %>% 
  filter(year == 2018)

mando_2019 <-  mando %>% 
  filter(year == 2019)

mando_2020 <-  mando %>% 
  filter(year == 2020)

mando_2021 <-  mando %>% 
  filter(year == 2021)
```


# LSPA Buoy Dissolved Oxygen Sensor History   

This script is designed to present the dissolved oxygen (DO) data from the Lake Sunapee Protective Association (LSPA)/Global Lake Ecological Observatory Network (GLEON) (Loon Island location) buoy (hereafter “the buoy”) alongside manual DO measurements made at the buoy and/or Longterm Monitoring Program (LMP) site 210, a monitoring site near the buoy location. This document is meant to serve as a reference document for colleagues who intend to use the dissolved oxygen data from the LSPA buoy. For the purposes of this document, only buoy data recorded at the Loon Island location are presented.  The data collection period covered is from mid-2007 until 2021. From deployment in mid-2007 until mid-2013, there was only a single __shallow__ (near-surface) DO senor in operation at the buoy, beginning in September 2013 a __deep__ sensor was added. 

__Calibration:__ The operation log states that the shallow DO unit was calibrated prior to moving the buoy to its Loon Island location in 2016, 2018 (HOBO U26 unit), 2019, 2020, and 2021. The deep DO sensor was added in September 2013, calibration was documented in 2016, 2019, 2020, and 2021. DO data are flagged with an ‘x’, indicating that there was no documentation of calibration, if there was no calibration at/before deployment of the buoy near Loon Island.  There may have been additional instances of calibration of these sensors, but they were not recorded in the operation log. 
This script is designed to present the dissolved oxygen (DO) data from the Lake Sunapee Protective Association (LSPA) buoy alongside manual DO measurements made at the buoy and/or site 210, a long term monitoring site near the buoy location. This is meant to serve as a reference document to other researchers who intend to use the dissolved oxygen data from the LSPA buoy. For the purposes of this document, only buoy data recorded at the Loon Island location are presented in this overview.

__Offsets:__ From 2013 through 2015, a number of offsets were introduced to one or both DO sensors in the Campbell datalogger program in an attempt to correct for the difference between manual DO measurements and the data recorded at the buoy. In this document, we have removed the offset and present the __'raw data'__, that is, data that were recorded without an offset or that were back-calculated from an offset value. Data handling of the offsets are detailed in the document *Sunapee-Buoy-Data-Logger-DO-Offset-2013-2015*. 

## Comparison Data

Because there are many methods for calibration and correction that will depend upon the end-user’s goals, here we provide the raw DO data as well as the manually collected DO data for users to choose their method of correction. The manually collected DO record is from both the neighboring LMP site 210 (data source: ‘LMP’, Steele et al. 2021) as well as DO readings taken at the buoy site (data source: ‘operations log’, unpublished data outside of this data package) as additional data that end-users may wish to use to apply correction factors to the data. The manually collected DO data are collated in the R script *buoy_manual_do_collation.r* and are stored in the file *manual_do_2007-2021_v2022-02-25.csv*.  

In this document, manual DO data plotted with the buoy’s shallow DO sensor were collected in the top 2m of the water, the manual measurements plotted with the buoy’s deep DO sensor were collected between 9 and 11m depth. The full record of LMP data is available in Steele et al. 2021.

Steele, B.G., Weathers, K. C., and LSPA. (2021). Quality controlled in situ data from multiple locations in Lake Sunapee, NH, USA from the Lake Sunapee Protective Association’s Long-term Monitoring Program, 1986-2020 (v2020.1) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.4652076

## Other LSPA Buoy data products

We note that there are other published data packages of the LSPA buoy data, for which corrections appropriate to the research questions asked in the study were applied. For example, the Richardson et al. 2020 data publication of the LSPA buoy data (EDI package ID 463) encompasses the years 2007-2013. The DO data in that package are from 2007-2012. The Richardson et al. data set was corrected for fouling and drift using a quadratic equation in an R script that is provided in the data package. This script references the manual DO measurements at site 210 as the calibration source and contains two methods for these types of corrections.

Richardson, D.C., C.C. Carey, K.C. Weathers, D.A. Bruesewitz, LSPA, and B.G. Steele. 2020. High Frequency Meteorological, Drift-Corrected Dissolved Oxygen, and Thermistor Temperature Data - Lake Sunapee Buoy, NH, USA, 2007 – 2013 ver 1 [Data set]. Environmental Data Initiative. https://doi.org/10.6073/pasta/846a36bf6fd704e508511e5f8a2ab3b5 

\newpage


## DO visual history

Below, we present a visual history for Lake Sunapee DO. We show DO concentration (mg/L) and DO saturation (percent) from the buoy sensors as well as from manual measurements obtained ~monthly by year. Manual measurements are made at site 210 (‘LMP’) are presented as gold symbols, manual measurements made at the buoy (‘operations log’) are light blue symbols, high-frequency data from the buoy are black symbols in all figures.

### Buoy flag definitions:


    x: no calibration on record

    o: data back-calculated from offset values

    wp: cleaning presumed (may not be visible at this level)
    
    w: sensor cleaned (may not be visible at this level)
    
    c: sensor calibrated (may not be visible at this level)
    
    i: data record is intermittent
    
    s: data are suspect
    
    sf: suspected sensor failure (2017 only)
    
    re: DO cap replacement, no calibration on record, presumed errant (2018 only)


### 2007 Shallow DO (1.5 meters):

```{r}
mando_2007_up <- mando_2007 %>% 
  filter(depth_m < 2) 

ggplot(do2007, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
   geom_point(data = mando_2007_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  geom_point() +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2007, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
   geom_point(data = mando_2007_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  geom_point() +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')

```

\newpage

### 2008 Shallow DO (1.5 meters):

```{r}
mando_2008_up <- mando_2008 %>% 
  filter(depth_m < 2) 

ggplot(do2008, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2008_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2008, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2008_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')

```

\newpage

### 2009 Shallow DO (1.5 meters):

```{r}
mando_2009_up <- mando_2009 %>% 
  filter(depth_m < 2) 

ggplot(do2009, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2009_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2009, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2009_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2010 Shallow DO (1.5 meters):

```{r}
mando_2010_up <- mando_2010 %>% 
  filter(depth_m < 2) 

ggplot(do2010, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2010_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2010, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2010_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2011 Shallow DO (1.5 meters):

```{r}
mando_2011_up <- mando_2011 %>% 
  filter(depth_m < 2) 

ggplot(do2011, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2011_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2011, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2011_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')

```

\newpage

### 2012 Shallow DO (1.5 meters):

```{r}
mando_2012_up <- mando_2012 %>% 
  filter(depth_m < 2) 

ggplot(do2012, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2012_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2012, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2012_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2013 Shallow DO (1.5 meters):

```{r}
mando_2013_up <- mando_2013 %>% 
  filter(depth_m < 2) 

ggplot(do2013, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2013_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2013, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2013_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')

```

\newpage

### 2013 Deep DO (10.5 meters):

```{r}
mando_2013_down <- mando_2013 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2013, aes(x = datetime, y = oxygenDissolved_mgl_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2013_down, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2013, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2013_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2014 Shallow DO (1.5 meters):

```{r}
mando_2014_up <- mando_2014 %>% 
  filter(depth_m < 2) 

ggplot(do2014, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2014_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2014, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2014_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')

```

\newpage

### 2014 Deep DO (10.5 meters):

```{r}
mando_2014_down <- mando_2014 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2014, aes(x = datetime, y = oxygenDissolved_mgl_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2014_down, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2014, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2014_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```
 

\newpage

### 2015 Shallow DO (1.5 meters):

```{r}
mando_2015_up <- mando_2015 %>% 
  filter(depth_m < 2) 

ggplot(do2015, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2015_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2015, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2015_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')

```

\newpage


### 2015 Deep DO (10.5 meters):

```{r}
mando_2015_down <- mando_2015 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2015, aes(x = datetime, y = oxygenDissolved_mgl_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2015_down, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2015, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2015_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2016 Shallow DO (1.5 meters):

```{r}
mando_2016_up <- mando_2016 %>% 
  filter(depth_m < 2) 

ggplot(do2016, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2016_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2016, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2016_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2016 Deep DO (10.5 meters):

```{r}
mando_2016_down <- mando_2016 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2016, aes(x = datetime, y = oxygenDissolved_mgl_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2016_down, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2016, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2016_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```


\newpage


### 2017 Shallow DO (1.5 meters):

```{r}
mando_2017_up <- mando_2017 %>% 
  filter(depth_m < 2) 

ggplot(do2017, aes(x = datetime, y = oxygenDissolved_mgl_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2017_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2017, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1p5m, shape = flag_do1p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2017_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2017 Deep DO (10.5 meters):

```{r}
mando_2017_down <- mando_2017 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2017, aes(x = datetime, y = oxygenDissolved_mgl_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2017_down, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2017, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2017_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```


\newpage

### 2018 Shallow DO (0.25 meters):

No shallow DO sensor data are available from buoy this season. Instead a HOBO U26 sensor deployed instead, note only DO concentration data available for this season.

```{r}
mando_2018_up <- mando_2018 %>% 
  filter(depth_m < 2) 

ggplot(do2018hobo, aes(x = datetime, y = oxygenDissolved_mgl_0p25m, shape = flag_do0p25m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2018_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

```


\newpage

### 2018 Deep DO (10.5 meters): 

```{r}
mando_2018_down <- mando_2018 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2018, aes(x = datetime, y = oxygenDissolved_mgl_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2018_down, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color =  data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2018, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10p5m, shape = flag_do10p5m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2018_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```


\newpage


### 2019 Shallow DO (0.25 meters):

```{r}
mando_2019_up <- mando_2019 %>% 
  filter(depth_m < 2) 

ggplot(do2019, aes(x = datetime, y = oxygenDissolved_mgl_0p25m, shape = flag_do0p25m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2019_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL, color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2019, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_0p25m, shape = flag_do0p25m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2019_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```



\newpage

### 2019 Deep DO (10 meters):

```{r}
mando_2019_down <- mando_2019 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2019, aes(x = datetime, y = oxygenDissolved_mgl_10m, shape = flag_do10m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2019_down, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2019, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10m, shape = flag_do10m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2019_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2020 Shallow DO (0.25 meters):

```{r}
mando_2020_up <- mando_2020 %>% 
  filter(depth_m < 2) 

ggplot(do2020, aes(x = datetime, y = oxygenDissolved_mgl_0p25m, shape = flag_do0p25m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2020_up, aes(x = datetime, y = oxygenDissolved_mgl,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2020, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_0p25m, shape = flag_do0p25m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2020_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,   color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2020 Deep DO (10 meters):

```{r}
mando_2020_down <- mando_2020 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2020, aes(x = datetime, y = oxygenDissolved_mgl_10m, shape = flag_do10m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2020_down, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2020, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10m, shape = flag_do10m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2020_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2021 Shallow DO (1 meter):

The DO data from the shallow sensor originate from the YSI EXO sensor. 

```{r}
mando_2021_up <- mando_2021 %>% 
  filter(depth_m < 2) 

ggplot(do2021upper, aes(x = datetime, y = oxygenDissolved_mgl_1m, shape = flag_exodo, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2021_up, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2021upper, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_1m, shape = flag_exodo, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2021_up, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct,shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```

\newpage

### 2021 Deep DO (10 meters):

```{r}
mando_2021_down <- mando_2021 %>% 
  filter(depth_m >= 9 & depth_m <= 11) 

ggplot(do2021lower, aes(x = datetime, y = oxygenDissolved_mgl_10m, shape = flag_do10m, color = data_source)) +
  geom_point() +
   geom_point(data = mando_2021_down, aes(x = datetime, y = oxygenDissolved_mgl, shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(mg/L)')

ggplot(do2021lower, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct_10m, shape = flag_do10m, color = data_source)) +
  geom_point() +
     geom_point(data = mando_2021_down, aes(x = datetime, y = oxygenDissolvedPercentOfSaturation_pct, shape = NULL,  color = data_source)) +
  facet_grid(buoy_location ~ .) +
  final_theme +
  scale_color_colorblind(name = 'data source\n(symbol color)') +
  scale_shape(name = 'DO flag\n(symbol shape)') +
  labs(x = NULL,
       y = 'dissolved oxygen\n(percent saturation)')
```
